Modeling and Simulation of Load Balancing Strategies for Computing in High Energy Physics

### 2018-08-17 paper/paper.tex
Add todo
Next iteration of paper

\newcommand{\TODO}[1]{\commentatside{#1}}
\usepackage[capitalise,nameinlink]{cleveref}
%Nice formats for \cref
\crefname{section}{Sect.}{Sect.}
\Crefname{section}{Section}{Sections}
\crefname{listing}{\lstlistingname}{\lstlistingname}
\Crefname{listing}{Listing}{Listings}
\usepackage{pgfplots}

@@

\author{Ren\'e Caspart${}^{1}$,
	Patrick Firnkes${}^{2}$,
	Manuel Giffels${}^{1}$,
	Anne Koziolek${}^{2}$,
	G\"unter Quast${}^{1}$,
	Ralf Reussner${}^{2}$ and 
	Maximilian Stemmer-Grabow${}^{2}$}

@@

Another challenge for the HEP community is the increasing heterogeneity and distribution of the computing resources of user jobs. Cloud resources, like HNSciCloud and AWS, or institute clusters and Tier 3 centers will be progressively introduced for the computation of these jobs. This results in a more complex scheduling, because each resource has a different optimal job types and mixtures. Furthermore, this leads to a more dynamic data placement, where fewer centers hosts data and thus the remote access is increasing.

@@

To solve both challenges, a more efficient usage of the grid and coping with the rising complexity of the system, a simulation of the WLCG should be created.
Currently, there is no model or simulation in existence to evaluate the load balancing strategies of the grid, helping the operators to make the correct decisions while optimizing the load balancing strategy.
\Cref{related} describes different attempts to simulate the WLCG or parts of it. However, they showed scalability issues or the simulations do not have the necessary detail level to simulate effects on working nodes that is necessary to evaluate load balancing strategies.

@@

The contribution of this paper is the modeling and simulating the performance of CMS computing jobs at the Tier 1 center GridKa as a proof of concept.
The model and simulation are created using the Palladio simulator. We extended this simulator to be able to simulate load balancing strategies. Furthermore, we created a automated model parameter analysis and model creation, to increase the usability.
This model enables us to evaluate different load balancing strategies for GridKa and enables us to draw conclusions about the best optimizations of the load balancing for the WLCG. Furthermore, it can be used as a foundation for models and simulations of larger parts of the WLCG.

This paper is organized as follows: \Cref{sec:palladio} presents the Palladio simulator as a foundation of the simulation. \Cref{sec:process} gives a overview on the simulation process. \Cref{sec:model} describes how different elements of the computing jobs and the computing resources are represented in our model and the implementation of the necessarily extension of Palladio. \Cref{sec:param} explains how we determine the model parameters and the automated model creation. The validation is presented in \Cref{validation} and \Cref{related} puts our work in context with the related work. In \Cref{outlook} possible design decisions and how they could be modeled and simulated are presented. Finally, \Cref{conlusion} gives a conclusion on this paper.

@@

Lehrig and Becker \cite{arch} extended Palladio with Architectural Templates to make it possible to model cloud computing environments efficiently.
Furthermore, Palladio's simulation approach SimuLizar was created to simulate self-adaptive systems \cite{becker2013simulizar}.
It was later extended to support the metrics scalability, elasticity, and efficiency \cite{arch}.

@@

\Cref{process} shows the simulation process starting with the matching of the data sets to receiving a valid model of the site and its jobs.

The starting point of the simulation process is to match the data sets from different data sources as a prerequisite  of the analysis. In our proof of concept we have three data sources: The CERN WMArchive, CMS GridKa job data and GridKa node performance data.

After matching and grouping the data for job types the analysis of the job and working node data to extract performance parameters for the model can take place. For example, these are the CPU demand of jobs, the frequency of the job types or the CPU power of the nodes. This analysis and the data set matching are automated and described in \Cref{sec:param}.

@@

After a job is scheduled on a resource container it acquires the specified number of job slots or waits if the node has not sufficient free slots. After that the CPU and I/O demand of the job are processed. To avoid one big I/O operation at the start of a job and make the processing of I/O more continuously, the processing is modeled using iterations that reflect the number of events. Thus, for each iteration there are I/O and CPU resources needed. Rather than simulating each event, we simulate a part of the processed events, because the simulation of each event would harm the scalability of the simulation. 
While processing, the job uses the specified number of cores to process the CPU demand. To simulate not explicitly modeled elements of the jobs such as the overhead created by pilot jobs we added a delay to each job. This delay is calibrated in order to match the measured throughput. After the job finished, it releases its acquired job slots. All of these parameters are specified as a probability function and are extracted from measured performance data as described in \Cref{sec:param}.

@@

As mentioned in \Cref{sec:model} a delay was added to the job to simulate not explicitly modeled elements. This delay was calibrated so that the measured throughput matches the simulated one of the calibration dataset. The simulated throughput of the test dataset matches the measured one accurately with only 1\% difference.
\Cref{shares} depicts the share of the simulated and measured job types, which also have a high accuracy.

\begin{figure}
	\centering
	\includegraphics[scale = 0.4]{images/shares}
	\caption[]{Measured and Simulated Shares of Job Types}
	\label{shares}
\end{figure}

@@

There are a variety of grid and cloud simulators existing, of which SimGrid and CloudSim are two popular ones. 
However, both of them are missing graphical model editors which makes the modeling of the computing jobs time-consuming \cite{simgrid}\cite{cloudsim}. The support of graphical editors is a big advantage of Palladio and increases its usability and efficiency. 
 parameterize

@@

Korenkov et. al simulated the JINR Tier1 site in 2016 \cite{jinr} and Hushchyn et al. created a simulation for the LHCb Grid in 2017 using SimGrid \cite{lhcb}. Both simulations focused on the effects of the location of the required job data on the throughput of the sites.

However, they neither modeled each working node nor modeled details of the scheduling process like job slots, thus both simulations would not be usable to simulate load balancing strategies on working node level.

\section{Outlook}
\label{outlook}
Our next steps are the evaluation of different design decisions. In this section a few design decisions are described that could be simulated.

@@

Also the effects of data caches could be simulated. Firstly, several sites must be modeled to simulate their interactions. Secondly, the jobs must be parametrized with the possibility of remote access. Then the data cache should be modeled. Finally, the performance improvements of the data cache by transferring data from the cache and not from sites can be simulated.
\section{Conclusion}
\label{conlusion}
We successfully simulated the CMS computing jobs at GridKa. To enable easy usage of the simulation we implemented an automated model parameter extraction and model creation. Furthermore, we extended Palladio to be able to simulate load balancing strategies. Our first results suggest that this approach could deliver accurate results regarding CPU efficiency, however it must be investigated what the reason for the data inconsistency and missing job information is. Regarding scalability we are confident that several computing centers could be simulated on a detailed level. For use cases that require the simulation of larger parts of the WLCG we suggest a more macroscopic model that focuses on the interaction of the sites and models the sites internals on a less detailed level.

This proof of concept showed that an automated model parameter extraction and model creation is possible and that Palladio can be used to evaluate load balancing strategies for computing in HEP. In contrast to the existing simulations our simulation models a site in a more detailed way, enabling the evaluation of load balancing strategies on working node level. Furthermore, we are confident that our approach helps to improve utilization of the grid and thus helps solving the challenges created by the HL-LHC update. The next step is the evaluation of a design decision mentioned in \Cref{outlook}.

\section*{References}
\bibliography{paper}

>>>>

\usepackage{todonotes}
\usepackage[capitalise,nameinlink]{cleveref}
%Nice formats for \cref
\crefname{section}{Sect.}{Sect.}
\Crefname{section}{Section}{Sections}
\crefname{listing}{\lstlistingname}{\lstlistingname}
\Crefname{listing}{Listing}{Listings}
\usepackage{pgfplots}

@@

\author{Ren\'e Caspart${}^{1}$,
	Patrick Firnkes${}^{2}$,
	Manuel Giffels${}^{1}$,
	Anne Koziolek${}^{2}$,
	G\"unter Quast${}^{1}$,
	Ralf Reussner${}^{2}$, 
	Maximilian Stemmer-Grabow${}^{2}$}

@@

Another challenge for the HEP community is the increasing heterogeneity and distribution of the computing resources of user jobs. Cloud resources, such as HNSciCloud and AWS, or institute clusters and Tier 3 centers will be progressively introduced for the computation of these jobs. This results in a more complex scheduling, because each resource has different optimal job types and mixtures. Furthermore, this leads to a more dynamic data placement, where fewer centers host data and thus remote access is increasing.

@@

To solve both challenges, a more efficient usage of the grid and coping with the rising complexity of the system, a simulation for the WLCG should be created.
Currently, there is no model or simulation in existence to evaluate the load balancing strategies of the grid, helping the operators to make the correct decisions while optimizing the load balancing strategy.
\Cref{related} describes different attempts to simulate the WLCG or parts of it. However, they showed scalability issues or the simulations do not have the necessary detail level to simulate effects on working nodes that is necessary to evaluate load balancing strategies.

@@

The contribution of this paper is the modeling and simulating of CMS computing jobs at the Tier 1 center GridKa as a proof of concept.
The model and simulation are created using the Palladio simulator. We extended this simulator to be able to simulate load balancing strategies. Furthermore, we created a automated model parameter analysis and model creation, to increase the usability.
This model enables us to evaluate different load balancing strategies for GridKa and enables us to draw conclusions about the best optimizations of the load balancing for the WLCG. Furthermore, it can be used as a foundation for models and simulations of larger parts of the WLCG.

This paper is organized as follows: \Cref{sec:palladio} presents the Palladio simulator as a foundation of the simulation. \Cref{sec:process} gives a overview on the simulation process. \Cref{sec:model} describes how different elements of the computing jobs and the computing resources are represented in our model. \Cref{sec:param} explains how we determine the model parameters and the automated model creation works. The validation is presented in \Cref{validation} and \Cref{related} puts our work in context with the related work. In \Cref{outlook} possible design decisions and how they could be modeled and simulated are presented. Finally, \Cref{conlusion} gives a conclusion on this paper.

@@

Lehrig and Becker \cite{arch} extended Palladio with Architectural Templates to make it possible to model cloud computing environments efficiently.
Furthermore, Palladio's simulation approach SimuLizar was created to simulate self-adaptive systems \cite{becker2013simulizar}.
It was later extended to support the metrics scalability, elasticity, and efficiency.

@@

\Cref{process} depicts the simulation process starting with matching of the data sets to receiving a valid model of the site and its jobs.

The starting point of the simulation process is matching the data sets from different data sources as a prerequisite  of the analysis. In our proof of concept we have three data sources: The CERN WMArchive, CMS GridKa job data and GridKa node performance data. These data sets have to be matched because each set contains different information about an executed job, but non of them contains the complete information about it.

After matching and grouping the data for job types, the analysis of the job and working node data to extract performance parameters for the model takes place. For example, these are the CPU demand of jobs, the frequency of the job types or the CPU power of the nodes. This analysis and the data set matching are automated and described in \Cref{sec:param}.

@@

After a job is scheduled on a resource container it acquires the specified number of job slots or waits if the node has not sufficient free slots. Afterwards the CPU and I/O demands of the job are processed. To avoid one big I/O operation at the start of a job and to make the processing of I/O more continuously, the processing is modeled using iterations that reflect the number of events. Thus, for each iteration there are I/O and CPU resources needed. Rather than simulating each event, we simulate a fraction of the processed events, because the simulation of each event would harm the scalability of the simulation. 
While processing, the job uses the specified number of cores to process the CPU demand. To simulate not explicitly modeled elements of the jobs such as the overhead created by pilot jobs we added a delay to each job. This delay is calibrated in order to match the measured throughput. After the job finished, it releases its acquired job slots. All of these parameters are specified as a probability function and are extracted from measured performance data as described in \Cref{sec:param}.

@@

As mentioned in \Cref{sec:model} a delay was added to the job to simulate the effects of not explicitly modeled elements. This delay was calibrated so that the measured throughput matches the simulated one of the calibration dataset. The simulated throughput of the test dataset matches the measured one accurately with only 1\% difference.
\Cref{shares} depicts the share of the simulated and measured job types, which also have a high accuracy.

\begin{figure}[h]
	\centering
	\includegraphics[scale = 0.6]{images/shares}
	\caption[]{Measured and Simulated Shares of Job Types}
	\label{shares}
\end{figure}

@@

There are a variety of grid and cloud simulators existing, of which SimGrid and CloudSim are two popular ones. 
However, both of them are implemented as libraries and not as standalone simulators. Thus, both are missing graphical model editors which makes the modeling of the computing jobs time-consuming \cite{simgrid}\cite{cloudsim}. The support of graphical editors is a big advantage of Palladio and increases its usability and efficiency. Another advantage of Palladio is the separation of the system into different models, which enables one to change on part of the system without modifying others. Furthermore, both do lack the support of an integrated visualization tool to display the results.

@@

Korenkov et. al simulated the JINR Tier1 site in 2016 \cite{jinr} and Hushchyn et al. created a simulation for the LHCb Grid in 2017 \cite{lhcb}. Both simulations focused on the effects of the location of the required job data on the throughput of the sites.
However, they neither modeled each working node nor modeled details of the scheduling process like job slots, thus both simulations would not be usable to simulate load balancing strategies on working node level.

\section{Outlook}
\label{outlook}
Our next steps are the evaluation of different design decisions. In this section a selection of possible design decisions that could be simulated are presented.

@@

Also the effects of data caches could be simulated. Firstly, several sites must be modeled to simulate their interactions. Secondly, the jobs must be parametrized with the possibility of remote access. Then the data cache should be modeled. Finally, the performance improvements of the data cache by transferring data from the cache and not from sites can be simulated.
\section{Conclusion}
\label{conlusion}

Computing resources for user analyses will become increasingly distributed and heterogeneous in the future and 
scheduling will be a key component for the efficient usage of the grid. We are confident that our approach helps to improve utilization of the grid and to be able to cope with the rising complexity of the system, thus serves the HEP community as a tool to solve the challenges created by the HL-LHC update.

We successfully simulated the CMS computing jobs at GridKa. To enable easy usage of the simulation we implemented an automated model parameter extraction and model creation. Furthermore, we extended Palladio to be able to simulate load balancing strategies. Our first results suggest that this approach could deliver accurate results regarding CPU efficiency, however it must be investigated what the reason for the data inconsistency and missing job information is. Regarding scalability we are confident that several computing centers could be simulated on a detailed level. For use cases that require the simulation of larger parts of the WLCG we suggest a more macroscopic model that focuses on the interaction of the sites and models the sites internals on a less detailed level.

This proof of concept showed that an automated model parameter extraction and model creation is possible and that Palladio can be used to evaluate load balancing strategies for computing in HEP. In contrast to the existing simulations our simulation models a site in a more detailed way, enabling the evaluation of load balancing strategies on working node level. The next step is the evaluation of design decisions mentioned in \Cref{outlook}.
\section*{References}
\bibliography{paper}
